\section*{Exercises}

\begin{exercise}
	Find the solution for the following linear system:
	$$
	\begin{array}{rcl}
	x & = & 4 \\
	y & = & 2x+1	
	\end{array}
	$$
	Is there one method for solving this system that is easier than the others? Give evidence for your answer. 
\end{exercise}
\bigskip

\begin{exercise}
	Find a linear system that has the solution $(2,5)$.
\end{exercise}
\bigskip

\begin{exercise}
Your philosophy teacher wants you to compare apples and oranges for an assignment. Your friend picks up a total of 12 bundles of fruit and paid \$74.00. The apples weighed 5 lbs per bundle and the oranges weighed 4 lbs per bundle. If the bundles of apples cost \$5 a piece and the bundles of oranges cost \$7 a piece, how many bundles of apples did your friend buy?
\end{exercise}

\begin{exercise}
Consider the linear system:
$$
\begin{array}{rcc}
y& = & x \\
y & = & -x 	
\end{array}
$$
with respective slopes 1 and -1. If we change the slopes of the two lines to 2 and -2, without changing the y-intercept, does it change the solution? If so, what is the new solution. Provide evidence for your claim. 	
\end{exercise}

\bigskip

\begin{exercise}
Solve the following system of equations using whatever method you choose:
$$
2x-y = 3
$$
$$
-3y = 4 +x
$$	
\end{exercise}

\bigskip

\begin{exercise}
	In what type of situations is it more difficult to use graphing than substitution or elimination to solve a linear system? Give an example of a system in which this is true. 
\end{exercise}

\bigskip

\begin{exercise}
Find the solution for the system:
$$
\begin{array}{rcl}
7x + 4y & = & 3 \\
x + \frac{1}{7}y & = & 3	
\end{array}
$$
\end{exercise}
\bigskip

\begin{exercise}
Solve the following system in two ways (using substitution and elimination) and compare your answers:
$$
\frac{1}{2}x + y = 6
$$	
$$
4x =y
$$

Should the solutions be the same or different. Provide evidence for your answer.
\end{exercise}

\bigskip

\begin{exercise}
It costs \$5.00 for a student and \$10.00 for non-students to see a movie at the Cinema 12. If the theater had 75 people go to a movie one night and made \$500 dollars from ticket sales, how many students saw a movie that night?
\end{exercise}

\bigskip

\begin{exercise}
	Find the solution for the linear system:
	$$
	2x+y=17
	$$
	$$
	\frac{1}{2}y=-x +1
	$$	
\end{exercise}

\begin{exercise}
Use whatever method you choose to find the solution for the system:
$$
\begin{array}{rcl}	
1002x + 404y &=  & 2016 \\ \\
4x + 6y &= &36
\end{array}
$$ 
\end{exercise}

\begin{exercise}
Solve the following linear system:
$$
2x+ y = 7 
$$
$$
\frac{1}{3}y = -\frac{2}{3}x + \frac{7}{3}
$$	
\end{exercise}
