

\section*{Exercises}

\begin{exercise}
	Given the following table, what is the average rate of change from 1 to 3?  Is this the same as the average rate of change from 2 to 4?  What about from 3 to 5?
	
	\begin{tabular}{|c|c|}
	\hline
	$x$ & $f(x)$ \\
	\hline
	0	&	-8\\
	\hline
	1	&	-1\\
	\hline
	2	&	2\\
	\hline
	3	&	6\\
	\hline
	4	&	10\\
	\hline
	5	&	13\\
	\hline
	\end{tabular}
\end{exercise}
\bigskip

\begin{exercise}
For the following four tables, guess whether they are going to be linear, and then graph the points.  If it is a linear function, draw a line through the points.  If it is not a linear function, draw whatever (safe for work) picture or curve you want that goes through all of the points.

	\begin{multicols}{4}
		\begin{tabular}{|c|c|}
		\hline
		$x$ & $f(x)$ \\
		\hline
		0	&	-2\\
		\hline
		1	&	-1.75\\
		\hline
		2	&	-1.5\\
		\hline
		3	&	-1.25\\
		\hline
		4	&	-1\\
		\hline
		5	&	-0.75\\
		\hline
		\end{tabular}
		
		\begin{tabular}{|c|c|}
		\hline
		$x$ & $f(x)$ \\
		\hline
		-10	&	-35\\
		\hline
		-5	&	45\\
		\hline
		0	&	75\\
		\hline
		5	&	55\\
		\hline
		10	&	-15\\
		\hline
		15	&	-145\\
		\hline
		\end{tabular}
		
		\begin{tabular}{|c|c|}
		\hline
		$x$ & $f(x)$ \\  %-11x - 6
		\hline
		0	&	-6\\
		\hline
		10	&	-116\\
		\hline
		5	&	-61\\
		\hline
		-1	&	5\\
		\hline
		12	&	-138\\
		\hline
		11	&	-127\\
		\hline
		\end{tabular}
		
		\begin{tabular}{|c|c|}
		\hline
		$x$ & $f(x)$ \\
		\hline
		0	&	4\\
		\hline
		1	&	4\\
		\hline
		2	&	4\\
		\hline
		3	&	4\\
		\hline
		4	&	4\\
		\hline
		5	&	4\\
		\hline
		\end{tabular}
	\end{multicols}
\end{exercise}
\bigskip

\begin{exercise}
The following table gives points for four linear functions, $f(x)$, $g(x)$, $h(x)$, and $j(x)$.  How many places do these four lines intersect each other?  How many of them are on this table?

	\begin{tabular}{|r|l|l|l|l|}
	\hline
	$x$ & $f(x)$ & $g(x)$ & $h(x)$ & $j(x)$\\
	\hline
	-5 & 3$\frac{2}{3}$ & -3 & 7 & 12\\
	\hline
	-4 & 3$\frac{1}{3}$ & -2$\frac{1}{2}$ & 5 & 10 \\
	\hline
	-3 & 3 & -2 & 3 & 8\\
	\hline
	-2 & 2$\frac{2}{3}$ & -1$\frac{1}{2}$ & 1 & 6\\
	\hline
	-1 & 2$\frac{1}{3}$ & -1 & -1 & 4\\
	\hline
	0 & 2 & -$\frac{1}{2}$ & -3 & 2\\
	\hline
	1 & 1$\frac{2}{3}$ & 0 & -5 & 0\\
	\hline
	2 & 1$\frac{1}{3}$ & $\frac{1}{2}$ & -7 & -2\\
	\hline
	3 & 1 & 1 & -9 & -4\\
	\hline
	4 & $\frac{2}{3}$ & 1$\frac{1}{2}$ & -11 & -6\\
	\hline
	5 & $\frac{1}{3}$ & 2 & -13 & -8\\
	\hline
	\end{tabular}
\end{exercise}
\bigskip

\begin{exercise}

\end{exercise}
\bigskip

\begin{exercise}
You're working for the WHO treating ebola in Sierra Leone.  Your facility treats children and adults, sending infants to a specialized facility.  One of your most important medications is for strengthening the heart and maintaining blood pressure.  Unfortunately, you only have 180000 units of medicine for the next three days.  Your goal is to fill as many of your 32 beds as you can, while using as much of your medication as possible, to maximize your resources.  You also have the following facts:

\begin{itemize}
\item Adults require 10 units per pound of body weight.
\item Children require 40 units per pound of body weight to account for smaller hearts.
\item The average child in the region weighs 55 pounds.
\item The average adult in the region weighs 140 pounds.
\end{itemize}

What is the largest number of children you will be able to take into your facility under these constraints?

\end{exercise}
\bigskip

\begin{exercise}
Because of your previous success, the administrators over your facility have called you to advise a decision about funding for a larger operation.  When treatments are successful and nearly complete, or when they are particularly unsuccessful, patients are transferred to another facility.  Furthermore, other hospitals will take note of success and send more patients to this one for treatment.

The administrators have picked up two linear trends relating funding to transfers.  They predict 3 daily transfers out of the facility for every \$8000 above \$60000 in its monthly budget.  They also predict a daily arrival for every \$4000 above \$49500.  They tell you that, if people are transferred out, they want the same number of people coming into the facility so that resources are being used optimally.  How much should they budget for the facility, monthly?

\end{exercise}
\bigskip

\begin{exercise}

\end{exercise}
\bigskip

\begin{exercise}
For the following function, give the horizontal and vertical intercepts:

$$f(x) = \frac{3}{4}x - \frac{1}{2}$$

For the following function, choose three inputs and find their corresponding outputs:

$$g(x) = -\frac{1}{3} + 1$$


\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
I started my party with a 24-pack of soda, and 2$\frac{1}{2}$ hours later, I had 10 left.  If we assume people are drinking at a constant rate, how much longer will the soda last?
\end{exercise}
\bigskip

\begin{exercise}
If you start out this month with an \$800 debt on your credit card, and if you make \$320 per week, give an equation for how much money you have as a function of time, assuming you have no other expenses.
\end{exercise}
\bigskip

\begin{exercise}
Write an equation of the line which passes through the points $(-4,2)$ and $(1,12)$.
\end{exercise}
\bigskip

\begin{exercise}
Write an equation of the line which passes through the points $(\frac{1}{2},\frac{1}{2})$ and $(1,2)$.
\end{exercise}
\bigskip

\begin{exercise}
A first line passes through the points $(1,4)$ and $(-1,-2)$.  Find the equation of a line perpendicular to this one which passes through the point $(3,3)$.
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip

\begin{exercise}
\end{exercise}
\bigskip