

\section*{Exercises}

\begin{exercise}
	You've just finished a party, and you're trying to combine all of your pizzas into as few boxes as possible for your refrigerator.  It's not important what types of pizza go together as long as they fit, and they're all the same size.  You have half of a cheese pizza, two thirds of a pepperoni pizza, eleven twelfths of a pineapple and anchovy pizza, one sixth of a vegetarian supreme pizza, and one quarter of a meatball pizza.
	
	How many boxes will all of this pizza take?
	
\end{exercise}
\bigskip

\begin{exercise}
	Write down the following as an equation, and then solve: four times the sum of a number and three is eight.
\end{exercise}
\bigskip

\begin{exercise}
	Write down the following as an equation, and then solve: one fourth of a number together with two and one half is five.
\end{exercise}
\bigskip

\begin{exercise}
	Write down the following as an equation, and then solve: eight minus a number is two squared.
\end{exercise}
\bigskip

\begin{exercise}
	Find an equivalent expression.  $3(x-y)$
\end{exercise}
\bigskip

\begin{exercise}
	Find an equivalent expression.  $2(2x - 2(x - 2))$
\end{exercise}
\bigskip

\begin{exercise}
	Simplify:
	
	$$\frac{2(3-1)-7}{18-(20-5)}$$
\end{exercise}
\bigskip

\begin{exercise}
	Simplify:
	
	$$\frac{4}{4(3+12)-3(10+9)}$$
\end{exercise}
\bigskip

\begin{exercise}
	Simplify:
	
	$$\frac{2(5+2)}{3} + \frac{4-5}{3(3-4)}$$
\end{exercise}
\bigskip

\begin{exercise}
A cookie recipe makes 18 cookies.  It calls for $\frac{3}{4}$ cup of butter.  If you only want to make 6 cookies, how many cups of butter will you need for the recipe?
\end{exercise}
\bigskip

\begin{exercise}
On your to-do list, you have tasks that take a quarter hour, half an hour, an hour and a half, and another half an hour.  How much time will it take to complete your to-do list?  Convert all of these times to minutes to check your answer.
\end{exercise}
\bigskip

\begin{exercise}
The sapling in my front yard grows about three quarters of an inch per month (30 days).  How much will it grow in 5 days?
\end{exercise}
\bigskip

\begin{exercise}
In a class of 40 students, there's enough pizza for each student to have three slices.  For the same amount of pizza, how many slices could each student have if only 24 showed up?
\end{exercise}
\bigskip

\begin{exercise}
Solve for $x$.

$2(x-1)+x = 3$
\end{exercise}
\bigskip

\begin{exercise}
Solve for $x$.

$10(5x+4) = -60$
\end{exercise}
\bigskip

\begin{exercise}
Solve for $x$.

$5x + 1 = -2(5-x)$
\end{exercise}
\bigskip

\begin{exercise}
You're deciding between two vehicles to buy based on the monthly cost for driving them.  Vehicle A gets 22 miles to the gallon, and insurance is \$60 per month.  Vehicle B gets 32 miles to the gallon, and insurance is \$90 per month.

If we assume gas is always \$2 per gallon, if you drive 500 miles per month, which vehicle will be cheaper?

How many miles would you have to drive before the monthly cost of the two vehicles is the same?
\end{exercise}
\bigskip

\begin{exercise}
You're considering getting a Netflix subscription.  You're primarily concerned with watching movies, so the main competitor for your purposes is Redbox.  A Netflix subscription is \$10 per month.  Each Redbox rental is \$1 per movie.

How many movies would you have to watch per month for Netflix to be the better deal?
\end{exercise}
\bigskip

\begin{exercise}
We have 45 meetings for this class this semester.  If we had some bad weather and had to miss 5 days of class, how much longer would we have to make each day so that you got your money's worth?
\end{exercise}
\bigskip

\begin{exercise}
You're writing four versions of a test as well as a quiz for your class.  Each test will have 10 multiple-choice questions, and the quiz will have one.  How many multiple-choice questions will you have to create?
\end{exercise}
\bigskip