\chapter{Introduction}



Hello. This is a fucking box.	




\begin{prblm}
This is a fucking boxed problem.	
\end{prblm}
	



\begin{prblm}
	Test?
\end{prblm}


\begin{pres_defn}[Farster]
Test drangulus. A farster is a powerful, yet extraordinarily weak, contradiction. It is always true, except every single situation.
	
\end{pres_defn}


\begin{defn}
Second definition. Joseph is an amateur.
\end{defn}

\begin{WrapBoxR}
This is the killer box that should be on the side of the text.	
\end{WrapBoxR}


yecdahiluzhnvurhnuanhurnhruinaufio c ur aiuor gaiu gyioa gria gri sr ya riia irg iayro gaior giagroiua iuor hauio riua ruio agriuoa guior oiau rguio aiu grauio gaiu graigaoiuegiau groiua goiur aoiudg. IYegiay doyi afiad iuah dygi agiyf gaiogr iagri ailur iuoaeiur ioiufr gaiogriu fgaiurg iosgri gsuirf hauior guialgruio agilruf agiulr guia rhuioariou oiw hrio aurh ophur.


\begin{WrapBoxL}
diai haur hruio ahdui aiuof ia hduio auior iul ahrui huiro auiro aior hiuoaiua h lkajhr ilr ailur ahru lhuri auil iau hruiruil aiuh rua iuluir uial iu ilh airui ailhruf iaru aigr auyz huila figravu9agliue fajdl hvuiaohuerila hruia iu gaio griul griorug aiugr gauio griua uipa.	
\end{WrapBoxL}



ilauhc aiouh fioaur aiu riourhui uar huila dia uhila rilu auirl aguirgaliuriaurailu valiug ioug aiulri aiuriul aiohui riualhuiua iu auiiuair gu gauil gi rgau agu faliu gdiul ailur gauil guia gfuirla gruila galiu grilua gilur giula agui gruil vziul rgvaui guzi gvuilz giu aliur zi uif gig vgyfzsgvg vs .

\pagebreak

\begin{WrapCurvedR}
asybiabvibaibdkfbaybiabe ukg y aryuk au gruagyru iuaegy avru gawe g airyiu fuha y.	
\end{WrapCurvedR}

Hrr auhr auyif sy ivaiyr iiar i airg ayg rgyaiu rgyiar gcsyrg iabriy graor gaygriuaygr ugfiuya geyagryuai gryiua gra guagfiya gryaui egaiurgaiugryuia iuryg gyrio aoiru gyior gfyaior gw.


\begin{pres_exmpl}
	If we take an egg and throw it at someones face, we get:
	$$
	4+7+2=2
	$$
	where 2 is not 2.
\end{pres_exmpl}