\chapter{Operations}



Numbers come in a lot of different forms. We all know that 2 is a number. Many of us have seen $\pi$ in some capacity and recognize that it is a number. In a similar manner, you may know that $\sqrt{9001}$ and $3i + 17$ are numbers. But without understanding what these numbers mean, they are worthless. The first step in our journey towards establishing meaning for numbers starts with classification of these numbers.

\begin{presentation}
\begin{defn}[Number Set]\label{defn1}
	A number set is a collection of numbers.
\end{defn}
\end{presentation} \index{Number Set}

First definition down. This is a small, but significant definition as well. Number sets will be a recurring character throughout this book. 

\begin{defn}[Natural Number]\label{defn2}
	A number is called natural if we can use it to count. We denote the set of natural numbers as $\mathbb{N}$. \\ $1,2,3,4,$ and $5$ are examples natural numbers.
\end{defn}

We use natural numbers every single day. However, there are some quantities that we cannot describe using natural numbers. Enter the integers.

\begin{presentation}
\begin{defn}[Integer]
	A number is an integer if it has no fractional part. An integer can be either positive or negative. We denote the set of integers as $\mathbb{Z}$. \\$-4, -2, 0, 3,17,  9002$ are all examples of integers. 
\end{defn}
\end{presentation}

We can use number lines to plot integers.

\vspace{2cm}

\numline









 Integers are also where we introduce mathematical operations. 



\begin{defn}[Operation]
An operation is a calculation that is done using some combination of numbers. The four elementary operations are addition (+), subtraction (-), multiplication ( $\cdot$ or $\times$), and division ($\div$ or $/$  ).
\end{defn}
The addition of two numbers is the combination of the two quantities. We can represent this operation on a number line.
\begin{example}[Addition on a Number Line]
	To visualize addition on a number line, start by plotting the first number on the number line. Consider $3+-6$.
	
	\begin{center}
	\begin{tikzpicture}
	\draw[latex-] (-6.5,0) -- (6.5,0) ;
	\draw[-latex] (-6.5,0) -- (6.5,0) ;
	\foreach \x in  {-5,-4,-3,-2,-1,0,1,2,3,4,5}
	\draw[shift={(\x,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
	\foreach \x in {-5,-4,-3,-2,-1,0,1,2,3,4,5 }
	\draw[shift={(\x,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
	{$\x$};
	\end{tikzpicture}
	\end{center}


	\begin{center}
	\begin{tikzpicture}
\begin{axis}[
  axis y line=none,
  axis lines=left,
  axis line style={-},
  xmin=-11,
  xmax=11,
  ymin=0,
  ymax=1,
  scatter/classes={o={mark=*}},
  restrict y to domain=0:1,
  xtick={-10,-9,-8, ...,10},
  width=16cm
]
\addplot table [y expr=0,meta index=1, header=false] {
114.06 o
119.94 o
};
\node[coordinate,label=above:{$L_1=114.06$}] at (axis cs:114.06,0.05) {};
\node[coordinate,label=above:{$L_2=119.94$}] at (axis cs:119.94,0.05) {};
\end{axis}
\end{tikzpicture}
	\end{center}

\end{example}




\begin{defn}[Fraction]
	A fraction is a number that expresses some part of an integer. It can also be viewed as an integer divided by an integer. 
\end{defn}

\section{Exercises}

\begin{enumerate}
\item 	
\end{enumerate}


\section{Problems}


\begin{enumerate}

\item Come up with a real world example for when adding or subtracting with negative numbers is necessary. 
\item 
\end{enumerate}


